---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)

# Remove trailing spaces from generated console output, including empty #> lines.
readme__output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
    output <- readme__output_hook(x, options)
    gsub("[ \\t]+(\\n|$)", "\\1", output, perl = TRUE)
})
```

# destep

<!-- badges: start -->
[![R-CMD-check](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`destep` is an R package for converting [DeST](https://www.dest.net.cn/)
building models to [EnergyPlus](https://energyplus.net/). It helps reuse
existing DeST models without rebuilding their inputs manually. Converted
models can be inspected, edited, and simulated with
[eplusr](https://cran.r-project.org/package=eplusr).

## Features

* Convert building geometry, materials, constructions, windows, doors, and
  exterior shading.
* Transfer schedules, internal heat and moisture gains, furniture heat storage,
  ventilation, and temperature and humidity controls.
* Generate load-calculation models and selected physical HVAC configurations.
* Convert DeST climate data to EnergyPlus EPW weather files.

`destep` is under active development. See `?to_idf` for supported model
features, conversion options, and EnergyPlus version requirements, and
[NEWS](NEWS.md) for recent changes.

## Installation

You can install the development version of destep like so:

```r
install.packages("destep",
    repos = c(
        hongyuanjia = "https://hongyuanjia.r-universe.dev",
        cran = "https://cran.r-project.org"
    )
)
```

## Get started

Window faces with the saved unbound marker `OF_ROOM = -1` are restored from
the matching side of their explicit host enclosure when that relationship is
unique and its room, outdoor or ground owner exists. Conflicting or ambiguous
bindings stop conversion. Only ownership and type are restored; geometry and
thermal inputs are retained. The default conversion works on a copy and leaves
the source database unchanged. Restorations produce a
`destep_restored_window_bindings` warning and are recorded in
`attr(idf, "conversion")$window_bindings` and saved IDF comments.

```{r example, message=FALSE, warning=FALSE, results='hide'}
library(destep)

# Download a DeST prototype model and load the target EnergyPlus dictionary
path <- download_dest_model("Commercial office A", "Chongqin", 2015, tempdir())
eplusr::use_idd("23.1", download = "auto")
```

```{r conversion, message=FALSE, warning=FALSE}
# Convert the model and its weather data
dest <- read_dest(path)
idf <- to_idf(dest, "23.1")
idf

epw <- to_epw(dest)
epw

DBI::dbDisconnect(dest)

# Save the EnergyPlus input and weather files
idf$save(file.path(tempdir(), "model.idf"))
epw$save(file.path(tempdir(), "weather.epw"))
```
