---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# destep

<!-- badges: start -->
[![R-CMD-check](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

> A toolkit to convert [DeST](https://www.dest.net.cn/) models to [EnergyPlus](https://energyplus.net/) models

## Installation

You can install the development version of destep like so:

```r
install.packages("destep",
    repos = c(
        hongyuanjia = "https://hongyuanjia.r-universe.dev",
        cran = "https://cran.r-project.org"
    )
)
```

## Supported conversion boundary

`destep` is under active development. The default conversion currently
preserves the following DeST components:

* Geometry, opaque constructions, doors, and windows
* Material thermal properties and aggregate or detailed glazing properties
* Hourly schedules from `SCHEDULE_YEAR`
* Outdoor ventilation from `ROOM_RELATION`
* Thermostat setpoints and Ideal Loads zone equipment from `ROOM_GROUP`
* Internal gains and occupant outdoor-air requirements
* Furniture storage from the effective room-type coefficient
* Ground temperatures, site metadata, and ground reflectance
* Exterior window overhangs and side fins
* Climate data through `to_epw()`

Furniture storage uses `ROOM.TYPE` to select
`ROOM_TYPE_DATA.FURNITURE_COEF`. The conversion follows the area-dependent
slab rule verified for DeST 0.2.230705, including its 50 m2 branch, and uses
EnergyPlus `InternalMass` with negligible radiative absorptance. Capacity
and free-floating temperature comparisons were checked independently in
EnergyPlus 26.1. This does not establish equivalence of DeST's inter-room
partition treatment or whole-building results.

For default DeST bshell runs, select
`to_eplus(..., people_heat = "temperature_dependent")` to include the
verified `max(0, input + 5.536 * (26 - previous temperature))` W/person rule.
The existing `"constant"` default corresponds to bshell's `--const_occupant`
option. The database does not identify that execution choice. The new mode
requires EnergyPlus 9.1 or newer, uses the previous zone-step temperature,
and preserves the source minimum occupancy and moisture input. Four separate
changing-occupancy tests passed in EnergyPlus 26.1; directional radiant
allocation and full-school agreement remain unresolved.

Lighting room heat includes `ROOM_TYPE_DATA.L_HEAT_RATE`: sensible heat is
the scheduled electrical power multiplied by this ratio. Lighting electricity
is preserved, while an unmetered sensible correction follows the same source
schedule and minimum power. This correction does not require EMS.

For the simplified DeST solar model, opt in to
`to_eplus(..., ver = "26.1", window_optics = "dest_solar")`. This maps SC
and two/three-pane counts to solar angle tables while preserving glass
resistance. The existing `"simple_glazing"` default remains available.
EnergyPlus retains solar-position, diffuse-integration and heat-balance
algorithms. This solar-only option does not preserve daylighting optics,
native glass storage, sky exchange or room solar recipient allocation.

Equipment moisture is read through `ROOM.TYPE` from
`ROOM_TYPE_DATA.E_MIN_HUM`, `E_MAX_HUM`, `E_PER_AREA`, and `E_SCHEDULE`.
The source rate is `minimum + (maximum - minimum) * schedule`, in kg/h,
multiplied by zone floor area for per-area inputs. Sensible equipment gains
remain independent. An all-latent `OtherEquipment` object with fuel type
`None` supplies moisture without adding an energy-meter entry. EMS adjusts
its latent watts using EnergyPlus's water-vapor enthalpy function and the
latest available zone temperature. Abrupt temperature changes can leave a
one-zone-step residual; whole-model DeST/EnergyPlus numerical equivalence
is not implied by this source mapping.
Nonzero equipment moisture requires EnergyPlus 9.1 or newer because older
versions lack the required EMS calling point. Use
`to_eplus(model, ver = "9.6.0", hvac = "ideal_loads")` for the real-model
path validated with this mapping. The current physical-HVAC path, which
targets 9.0.1, still rejects nonzero equipment moisture.

The default `hvac = "ideal_loads"` path is not restricted to a particular
air-system topology or to one or two conditioned rooms. It creates a
`ZoneHVAC:IdealLoadsAirSystem` for each convertible air-conditioned DeST room
and maps its availability, thermostat, humidity-control, and outdoor-air
inputs. When a room group references a DeST air-conditioning system, this path
does not reconstruct its physical fans, coils, ducts, air loop, or plant.

The opt-in `hvac = "physical"` path groups conditioned rooms by their referenced
`AC_SYS` and generates each supported air loop independently. It accepts:

* one-room `AC_SYS_TYPE = 0` constant-volume systems;
* `AC_SYS_TYPE = 0` systems with multiple rooms, represented by
  constant-volume fans and terminal-reheat branches; and
* `AC_SYS_TYPE = 1` systems with one or more rooms, represented by VAV
  terminal-reheat branches.

A model may contain multiple supported `AC_SYS` records. Each system must have
one `AHU`; rooms in a multizone system must share one availability schedule;
and supported systems may use `FRESH_AIR_TYPE` 1, 5, or 6. Parameters absent
from the DeST model, including allocations named for every room served by a
terminal-reheat path, must be supplied through `hvac_options`; see `?to_eplus`
for the required fields. The current flat option list applies the same equipment
assumptions to every generated air loop. All generated air loops share one
option-defined chilled-water plant and, when required, one hot-water plant.

Existing real-model numerical regression coverage comprises a one-room type-0
system, a two-room type-0 system, and a two-room type-1 system. Higher room
counts and multiple-system assembly are implemented, but remain outside that
numerical evidence set until suitable real source models and reference results
are added. Other physical topologies are not projected. Source models containing
an `AC_SYS_TYPE` other than 0 or 1 stop with an explicit error in either HVAC
mode.

## EnergyPlus versions

The reproducible conversion baseline is
[EnergyPlus](https://energyplus.net/) 9.0.1. The default load-only conversion
can request versions supported by
[eplusr](https://cran.r-project.org/package=eplusr), with compatibility
warnings where a version-specific behavior has not been verified. The
physical HVAC paths currently require EnergyPlus 9.0.1. A converted 9.0.1
model can be transitioned to a newer EnergyPlus version with `eplusr`.
Aggregate `WindowMaterial:SimpleGlazingSystem` output targeting EnergyPlus
9.0 through 9.3 emits a warning because EnergyPlus corrected the relevant
angular-reflectance behavior in version 9.4.

## Get started

```{r example, message=FALSE}
library(destep)

# use a DeST typical building model as an example
path <- download_dest_model("Commercial office A", "Chongqin", 2015, tempdir())

# make sure EnergyPlus IDD file can be found even if EnergyPlus itself was not
# installed
eplusr::use_idd("9.0.1", download = "auto")

# read once, then create the EnergyPlus input and weather objects
dest <- read_dest(path)
idf <- to_eplus(dest, "9.0.1")
epw <- to_epw(dest)

output_dir <- file.path(tempdir(), "destep-readme")
dir.create(output_dir)
idf$save(file.path(output_dir, "model.idf"))
epw$save(file.path(output_dir, "weather.epw"))
DBI::dbDisconnect(dest)

file.exists(file.path(output_dir, c("model.idf", "weather.epw")))
```
