---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# destep

<!-- badges: start -->
[![R-CMD-check](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

> A toolkit to convert [DeST](https://www.dest.net.cn/) models to [EnergyPlus](https://energyplus.net/) models

## Installation

You can install the development version of destep like so:

```r
install.packages("destep",
    repos = c(
        hongyuanjia = "https://hongyuanjia.r-universe.dev",
        cran = "https://cran.r-project.org"
    )
)
```

## Supported conversion boundary

`destep` is under active development. The default conversion currently
preserves the following DeST components:

* Geometry, opaque constructions, doors, and windows
* Material thermal properties and aggregate or detailed glazing properties
* Hourly schedules from `SCHEDULE_YEAR`
* Outdoor ventilation from `ROOM_RELATION`
* Thermostat setpoints and Ideal Loads zone equipment from `ROOM_GROUP`
* Internal gains and occupant outdoor-air requirements
* Ground temperatures, site metadata, and ground reflectance
* Exterior window overhangs and side fins
* Climate data through `to_epw()`

The default `hvac = "ideal_loads"` path is not restricted to a particular
air-system topology or to one or two conditioned rooms. It creates a
`ZoneHVAC:IdealLoadsAirSystem` for each convertible air-conditioned DeST room
and maps its availability, thermostat, humidity-control, and outdoor-air
inputs. When a room group references a DeST air-conditioning system, this path
does not reconstruct its physical fans, coils, ducts, air loop, or plant.

The opt-in `hvac = "physical"` path currently implements three explicitly
validated model structures:

* one conditioned room connected to one `AC_SYS_TYPE = 0` constant-volume
  system;
* exactly two conditioned rooms connected to one shared `AC_SYS_TYPE = 0`
  constant-volume terminal-reheat system; and
* exactly two conditioned rooms connected to one shared `AC_SYS_TYPE = 1` VAV
  terminal-reheat system.

These are the current boundaries of direct physical-system generation, not the
boundaries of the default load-oriented conversion. Supported physical systems
may use `FRESH_AIR_TYPE` 1, 5, or 6. Equipment and zone outdoor-air parameters
that are absent from the DeST model must be supplied through `hvac_options`;
see `?to_eplus` for the required fields. Other physical topologies are not yet
projected. At present, source models containing an `AC_SYS_TYPE` other than 0
or 1 stop with an explicit error in either HVAC mode.

## EnergyPlus versions

The reproducible conversion baseline is
[EnergyPlus](https://energyplus.net/) 9.0.1. The default load-only conversion
can request versions supported by
[eplusr](https://cran.r-project.org/package=eplusr), with compatibility
warnings where a version-specific behavior has not been verified. The
physical HVAC paths currently require EnergyPlus 9.0.1. A converted 9.0.1
model can be transitioned to a newer EnergyPlus version with `eplusr`.
Aggregate `WindowMaterial:SimpleGlazingSystem` output targeting EnergyPlus
9.0 through 9.3 emits a warning because EnergyPlus corrected the relevant
angular-reflectance behavior in version 9.4.

## Get started

```{r example, message=FALSE}
library(destep)

# use a DeST typical building model as an example
path <- download_dest_model("Commercial office A", "Chongqin", 2015, tempdir())

# make sure EnergyPlus IDD file can be found even if EnergyPlus itself was not
# installed
eplusr::use_idd("9.0.1", download = "auto")

# read once, then create the EnergyPlus input and weather objects
dest <- read_dest(path)
idf <- to_eplus(dest, "9.0.1")
epw <- to_epw(dest)

output_dir <- file.path(tempdir(), "destep-readme")
dir.create(output_dir)
idf$save(file.path(output_dir, "model.idf"))
epw$save(file.path(output_dir, "weather.epw"))
DBI::dbDisconnect(dest)

file.exists(file.path(output_dir, c("model.idf", "weather.epw")))
```
