---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# destep

<!-- badges: start -->
[![R-CMD-check](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

> A toolkit to convert 'DeST' models to 'EnergyPlus' models

## Installation

You can install the development version of destep like so:

```r
install.packages("destep",
    repos = c(
        hongyuanjia = "https://hongyuanjia.r-universe.dev",
        cran = "https://cran.r-project.org"
    )
)
```

## Supported DeST components

This package is still under heavy development and is not ready for use. Currently, the following components are supported:

* [x] Geometry
* [x] Material
* [x] Construction
* [ ] Internal gains [WIP]
* [ ] Shading
* [ ] HVAC

## Get started

```{r example}
library(destep)

# use a DeST typical building model as an example
path <- download_dest_model("Commercial office A", "Chongqin", 2015, tempdir())

# make sure EnergyPlus IDD file can be found even if EnergyPlus itself was not
# installed
eplusr::use_idd(23.1, download = "auto")

# read the DeST model and convert it to an EnergyPlus model
read_dest(path) |> to_eplus(23.1)
```
