---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# destep

<!-- badges: start -->
[![R-CMD-check](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hongyuanjia/destep/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`destep` is an R package for converting [DeST](https://www.dest.net.cn/)
building models to [EnergyPlus](https://energyplus.net/). It helps reuse
existing DeST models without rebuilding their inputs manually. Converted
models can be inspected, edited, and simulated with
[eplusr](https://cran.r-project.org/package=eplusr).

## Features

* Convert building geometry, materials, constructions, windows, doors, and
  exterior shading.
* Transfer schedules, internal heat and moisture gains, furniture heat storage,
  ventilation, and temperature and humidity controls.
* Generate load-calculation models and selected physical HVAC configurations.
* Convert DeST climate data to EnergyPlus EPW weather files.

`destep` is under active development. See `?to_eplus` for supported model
features, conversion options, and EnergyPlus version requirements, and
[NEWS](NEWS.md) for recent changes.

## Installation

You can install the development version of destep like so:

```r
install.packages("destep",
    repos = c(
        hongyuanjia = "https://hongyuanjia.r-universe.dev",
        cran = "https://cran.r-project.org"
    )
)
```

## Get started

```{r example, message=FALSE, warning=FALSE, results='hide'}
library(destep)

# Download a DeST prototype model and load the target EnergyPlus dictionary
path <- download_dest_model("Commercial office A", "Chongqin", 2015, tempdir())
eplusr::use_idd("23.1", download = "auto")

# Convert the model and its weather data
dest <- read_dest(path)
idf <- to_eplus(dest, "23.1")
epw <- to_epw(dest)
DBI::dbDisconnect(dest)

# Save the EnergyPlus input and weather files
idf$save(file.path(tempdir(), "model.idf"))
epw$save(file.path(tempdir(), "weather.epw"))
```
